\chapter{Background}
This section aims to introduce the reader to concepts relevant to the Flow DSL. This is split into four sections: High Level Synthesis (\ref{hls}), the Dataflow Programming Model (\ref{dataflow}), Embedded Domain-Specific Languages (\ref{edsl}) and Intermediate Representation Frameworks (\ref{irf}).

\input{incl/23_hls.tex}

\input{incl/26_dataflow.tex}

\input{incl/28_dsl.tex}

\input{incl/29_mlir.tex}

\section{Ethical Considerations}
At this project is on developing a system design tool, there is a non-zero chance, as with any freely available tools in this field, that the project may be used in military applications by any state or by bad actors to produce hardware with the intent of subverting organisations and governments.

As attempting to introduce measures to prevent military use would likely have a detrimental effect on the functionality of the tool for regular users, and bad actors using freely available software to commit crimes does not change the illegality of these crimes.

When considering the direct ethical impacts of the tools created and presented in this paper, there is nothing notable to report as no personal data is being used and using more efficient hardware will reduce power consumption and therefore greenhouse gas emissions.

It is also useful to note that all code used in this project will be released on GitHub at the conclusion of this project in the \href{https://github.com/1rre/FlowDSL}{1rre/FlowDSL} repository. This will provide further resources to those who wish to use xDSL for hardware or to investigate static streaming dataflow programming for hardware.

\par\noindent\hrulefill\par

In summary, this chapter has provided the foundational knowledge required to understand the theoretical underpinnings of Flow DSL, with regards to High Level Synthesis, the Dataflow Programming Model, Embedded Domain-Specific Languages, and Intermediate Representation Frameworks. Together, these areas encompass the main components and concepts that form the basis of Flow DSL and its operation.

We also delved into the ethical implications of this work, acknowledging the potential for misuse while emphasizing the project's primary purpose as a system design tool meant to enhance productivity and efficiency in hardware design.